Discriminative Sensing Based on Signal Processing for Information Security Analysis

### 2017-11-27 docs/thesis/slides/apresentacao_unb.tex
eigen similarity

	This thesis outlines the development and evaluation of approaches based on \textbf{signal processing for information security analysis}

@@

		\item To make the data discriminative;
		\item For network attack detection, mobile malicious behavior analysis and fraud detection in mobile money transactions.

@@

		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
		\item Eigen similarity analysis for extracting detailed information about accurate time and network ports under attack;
		\item Accuracy and performance evaluation through experimental scenario and the DARPA 1998 dataset.

@@

	This thesis also proposes an \textbf{architecture and approach based on user behavior analysis of mobile apps}

@@

	We also propose a \textbf{tensor-based dictionary learning method for fraud detection from imbalanced data}, in order to apply the tensor decomposition for dictionary learning methods to evidentiate the \textbf{discriminative sensing of a fraud detection dataset}

@@

		\item A sparse representation based classification (SRC) method through learning a tensor-based dictionary;

@@

		\item Evaluate the reconstruction error for fraud classification of mobile money transactions.

@@

\begin{frame}{Contributions}
	\begin{enumerate}
		\item An approach based on \textbf{eigen similarity analysis} for extracting detailed information about accurate time and network ports under network attack;
		\item Computational \textbf{complexity analysis} of the proposed framework;
		\item An architecture and techniques for \textbf{offline behavioral analysis} of a corporate mobile client security architecture;
		\item A \textbf{tensor-based dictionary learning} approach for \textbf{fraud detection} in mobile payment transactions;
	\end{enumerate}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	SECTION: Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Introduction
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Introduction}
	We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise})
	\begin{itemize}
		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
		\item Eigen similarity analysis for extracting detailed time and network ports under attack. 
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Introduction
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Introduction}
	We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise})
	\begin{itemize}
		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
		\item Eigen similarity analysis for extracting detailed time and network ports under attack. 
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Related works
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Related works}

	Classical methods typically employ data mining [39, 34, 70] and regular file analysis [76] to detect patterns that indicate the presence of specific attacks in network traffic;

	Callegari \emph{et al} [104] propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data. However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.

@@

	Lee \emph{et al.} [61] propose OverSampling PCA (osPCA), determines the anomaly of the target according to the variation of eigenvector by similarity analysis and over sampling. In contrast, we apply MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack, for flood and probe attacks;

	Lu and Ghorbani [63] proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory. However, their work requires a training method and presents limitations on identification of behaviors without significant outliers;

	Zonglin \emph{et al} [104] proposed a method to detect traffic anomaly with correlation analysis, but the work is not applied to probe and flood attack detection.

	Focus on \textbf{probe} and \textbf{flooding} attacks;

	Two scenarios: A \textbf{synthetic dataset} as a signal superposition of legitimate traffic, noise and malicious traffic, and selected cases of \textbf{DARPA dataset} that reproduce probe and flooding attacks.

@@

	7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.

	The attacks in this dataset can be grouped into: denial-of-service (DoS); remote to local (R2L); user to root (U2R); and probe attack.

	The most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.

	We select the cases that reproduce probe and flooding attacks.

@@

	For flooding detection, calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
	\begin{equation}\label{eq:eq02}
		\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
	\end{equation}

@@

	For probing detection, compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
	\begin{equation}\label{eq:eq04}
		\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
	\end{equation}

@@

	Compute eigenvalue decomposition (EVD) to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
	\begin{equation}\label{eq:eq06}
	\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
	\end{equation}
	\begin{equation}\label{eq:eq060}
	\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
	\end{equation}

@@

	\begin{equation}\label{eq:eq07}
		\boldsymbol{E} =
		\begin{bmatrix}
			\lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
			\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
			\vdots & \vdots & \ddots & \vdots  \\
			\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
		\end{bmatrix}.
	\end{equation}

@@

	Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.

	The proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame to identify variations and estimate the model order $\hat{d}$ (estimated number of time frames under attack)

	$\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to

	$\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigen Similarity Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

@@

	$\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack

	$\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the flood and port scan attack identification.

	The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack. Each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
	\begin{equation}\label{eq:eq12}
		\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
	\end{equation}

@@

	We evaluate three approaches for eigen similarity analysis: 
	\begin{itemize}
		\item \textbf{incremental:} concatenates eigenvectors incrementally;
		\item \textbf{individual:} concatenates each eigenvector individually and discards;
		\item \textbf{incremental individualized:} incremental until detect attach and individual after.
	\end{itemize}

@@

	For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack is compared against the $\boldsymbol{v}_{(n)}$ identified as under attack, individually evaluating the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.

@@

	\begin{table}[h!]
	  \centering
	  \scriptsize
	  \caption{Largest Eigenvalue related to attacks detection}
	  \label{tab:tab3}
	  \begin{tabular}{ c c c c c }
		\toprule
		\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
				\hhline{~----}
			&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
			&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
		\midrule
		1 &1887545 &1887545 &1887545 &2,0734 \\
		2 &2341327 &2341327 &2341327 &2,1451 \\
		3 &3213867 &3213867 &3213867 &10,0718 \\
		4 &133238294 &133238294 &731229 &2,1620 \\
		5 &92384021611 &6367983 &92384021611 &2,4253 \\
		6 &708335 &708335 &708335 &1,7948 \\
	    \bottomrule
	  \end{tabular}
	\end{table}

@@

	\begin{table}[h!]
	  \centering
	  % \scriptsize
	  \caption{MOS schemes applied to port scan and flood detection}
	  \label{tab:tab4}
	  \begin{tabular}{ c c c c c c c c }
		\toprule
		\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
				\hhline{~------~}
			&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
		\midrule
		Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
		Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
		\midrule
		Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
		Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
		\midrule
		Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
		Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
		\midrule
		Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
		Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
	    \bottomrule
	  \end{tabular}
	\end{table}

@@

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Port Scan Detection}
  \label{tab:tab5}
  \begin{tabular}{ c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~~---~}
			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
	\midrule
	3 &1 &0.9946 &0.9946 &0.9946 &no \\
	3 &2 &0.9934 &0.9934 &0.9999 &no \\
	3 &3 &0.9912 &0.9912 &0.9999 &no \\
	3 &4 &0.9888 &0.9888 &0.9999 &no \\
	3 &5 &0.9856 &0.9856 &0.9998 &no \\
	3 &6 &0.9840 &0.9840 &0.9999 &no \\
	3 &7 &0.9824 &0.9824 &1.0000 &no \\
	3 &8 &0.9794 &0.9794 &0.9999 &no \\
	3 &9 &0.9673 &0.9673 &0.9926 &no \\
	3 &10 &0.9674 &0.9674 &0.9997 &no \\
	3 &11 &0.9733 &0.9733 &0.9993 &no \\
	3 &12 &0.9702 &0.9702 &0.9993 &no \\
	3 &13 &0.9677 &0.9677 &0.9999 &no \\
	3 &14 &0.9646 &0.9646 &0.9998 &no \\
	3 &15 &0.0216 &0.0216 &0.0276 &yes \\
	3 &16 &0.9621 &0.0209 &1.0000 &no \\
	3 &17 &0.9611 &0.0199 &0.9998 &no \\
	3 &18 &0.9612 &0.0191 &0.9999 &no \\
	3 &19 &0.9613 &0.0186 &0.9998 &no \\
	3 &20 &0.9638 &0.0190 &1.0000 &no \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Synflood Detection}
  \label{tab:tab6}
  \begin{tabular}{ c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~~---~}
			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
	\midrule
	4 &1 &1.0000 &1.0000 &1.0000 &no \\
	4 &2 &0.9999 &0.9999 &1.0000 &no \\
	4 &3 &0.9997 &0.9997 &0.9999 &no \\
	4 &4 &0.9998 &0.9998 &1.0000 &no \\
	4 &5 &0.9965 &0.9965 &0.9908 &no \\
	4 &6 &0.9975 &0.9975 &1.0000 &no \\
	4 &7 &0.9977 &0.9977 &1.0000 &no \\
	4 &8 &0.9980 &0.9980 &1.0000 &no \\
	4 &9 &0.9987 &0.9987 &0.9999 &no \\
	4 &10 &0.9991 &0.9991 &1.0000 &no \\
	4 &11 &0.0085 &0.0085 &0.0284 &yes \\
	4 &12 &0.0162 &0.0120 &0.0343 &yes \\
	4 &13 &0.0248 &0.0158 &0.0427 &yes \\
	4 &14 &0.1243 &0.0185 &0.1041 &yes \\
	4 &15 &0.0082 &0.0162 &0.0103 &yes \\
	4 &16 &0.0404 &0.0070 &0.0580 &yes \\
	4 &17 &0.0397 &0.0007 &0.0573 &yes \\
	4 &18 &0.0408 &0.0042 &0.0584 &yes \\
	4 &19 &0.0408 &0.0079 &0.0584 &yes \\
	4 &20 &0.0477 &0.0092 &0.0757 &yes \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[h!]
  \centering
  % \footnotesize
  \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
  \label{tab:tab9}
  \begin{tabular}{ c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~--~}
			&\textbf{Incremental Individualized} &\textbf{Individual}\\
	\midrule
	80 &1.0000 &1.0000 &no \\
	443 &1.0000 &1.0000 &no \\
	53 &1.0000 &1.0000 &no \\
	21 &1.0000 &1.0000 &nos \\
	22 &1.0000 &1.0000 &no \\
	23 &1.0000 &1.0000 &no \\
	25 &1.0000 &1.0000 &no \\
	110 &1.0000 &1.0000 &no \\
	143 &1.0000 &1.0000 &no \\
	161 &1.0000 &1.0000 &no \\
	69 &1.0000 &1.0000 &no \\
	123 &1.0000 &1.0000 &no \\
	445 &1.0000 &1.0000 &no \\
	600 &0.0077 &0.0427 &yes \\
	19 &1.0000 &1.0000 &no \\
	67 &1.0000 &1.0000 &no \\
	68 &1.0000 &1.0000 &no \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

	\textbf{This thesis} outlines the development and evaluation of approaches based on \textbf{signal processing for information security analysis}...

@@

		\item To make the data \textbf{discriminative} for network attack detection, mobile malicious behavior analysis and fraud detection in mobile money transactions.

@@

		\item MOS and eigenvalue analysis are applied to network attack detection;
		\item Eigen similarity analysis identify accurate time and network ports under attack;
		\item Evaluation through experimental scenario and the DARPA 1998 dataset.

@@

	This thesis also proposes an \textbf{approach malicious behavior detection in mobile apps}

@@

	We also propose a \textbf{tensor-based dictionary learning method for fraud detection from imbalanced data} 

@@

		\item To evidentiate the \textbf{discriminative sensing of a fraud detection dataset};
		\item Sparse representation based classification (SRC) through learning a tensor-based dictionary;

@@

		\item Classify fraud classification of mobile money transactions according to the reconstruction error.

@@

% \begin{frame}{Contributions}
% 	\begin{enumerate}
% 		\item An approach based on \textbf{eigen similarity analysis} for extracting detailed information about accurate time and network ports under network attack;
% 		\item Computational \textbf{complexity analysis} of the proposed framework;
% 		\item An architecture and techniques for \textbf{offline behavioral analysis} of a corporate mobile client security architecture;
% 		\item A \textbf{tensor-based dictionary learning} approach for \textbf{fraud detection} in mobile payment transactions;
% 	\end{enumerate}
% \end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	SECTION: Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Introduction
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Introduction}
	\begin{itemize}
		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
		\item We conduct a performance evaluation for complexity analysis and processing time.
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Related works
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Related works}

	\begin{itemize}
		\item Classical methods typically employ data mining [39, 34, 70] and regular file analysis [76] to detect patterns that indicate the presence of specific attacks in network traffic;
		\item Callegari \emph{et al} [104] propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data. However, Callegari \emph{et al} \textbf{focus on flood attack detection}, not addressing probe attack detection, and their approach relies on \textbf{visual analysis}.
	\end{itemize}

@@

	\begin{itemize}
		\item Lee \emph{et al.} [61] propose OverSampling PCA, determines the \textbf{anomaly} according to the variation of \textbf{eigenvector by similarity analysis and over sampling}. We apply \textbf{MOS for attack detection} and \textbf{similarity analysis for identification} of time and ports under attack;
		\item Lu and Ghorbani [63] proposed a \textbf{network anomaly} detection model based on \textbf{network flow, wavelet approximation, and system identification theory}. However, their work requires a \textbf{training} and presents \textbf{limitations} on identification of behaviors without \textbf{significant outliers}.
	\end{itemize}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Datasets
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Datasets}

	We focus on \textbf{probe} and \textbf{flooding} attacks;

	Two scenarios: 
	\begin{itemize}
		\item A \textbf{synthetic dataset} as a signal superposition of legitimate traffic, noise and malicious traffic;
		\item Selected cases of \textbf{DARPA dataset} that reproduce probe and flooding attacks.
	\end{itemize}

@@

	\begin{itemize}
		\item 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks;
		\item The most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack;
		\item We select the cases that reproduce probe and flooding attacks.
	\end{itemize}

@@

	For \textbf{flooding} detection, calculate the \textbf{sample covariance matrix} $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
	\begin{equation}\label{eq:eq02}
		\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
	\end{equation}

@@

	For \textbf{probing} detection, compute the \textbf{sample covariance} $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have \textbf{zero mean and unitary standard deviation} as follows
	\begin{equation}\label{eq:eq04}
		\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
	\end{equation}

@@

	Compute eigenvalue decomposition (EVD) to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix (time frame $q$), according to (\ref{eq:eq060}).
	\begin{equation}\label{eq:eq06}
	\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
	\end{equation}
	\begin{equation}\label{eq:eq060}
	\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
	\end{equation}

@@

	\begin{equation}\label{eq:eq07}
		\boldsymbol{E} =
		\begin{bmatrix}
			\color{red}\lambda_1^{(1)} & \color{red}\lambda_1^{(2)} & \color{red}\lambda_1^{(3)} & \cdots & \color{red}\lambda_1^{(Q)} \\
			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
			\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
			\vdots & \vdots & \ddots & \vdots  \\
			\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
		\end{bmatrix}.
	\end{equation}

@@

	\begin{itemize}
		\item Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$;
		\item $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
		\item $\hat{d}$ is the estimated number of time frames under attack.
	\end{itemize}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigenvalue Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigenvalue Analysis}

	Through eigenvalue analysis, we identify the time frames $q$ under attack

@@

	The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack.

	Each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$
	\begin{equation}\label{eq:eq12}
		\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
	\end{equation}

@@

	$\boldsymbol{v}^{(q)}$ is the most significant eigenvectors

	$\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of one network traffic

	We evaluate \textbf{three approaches} for eigen similarity analysis: 
	\begin{itemize}
		\item \textbf{incremental:} concatenates eigenvectors incrementally;
		\item \textbf{individual:} concatenates each eigenvector individually and discards;
		\item \textbf{incremental individualized:} incremental until detect attach and individual after.
	\end{itemize}

@@

	For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack is compared against the $\boldsymbol{v}_{(n)}$ identified as under attack,

	Evaluate the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.

@@

	\begin{table}[h!]
	  \centering
	  \scriptsize
	  \caption{Largest Eigenvalue related to attacks detection}
	  \label{tab:tab3}
	  \begin{tabular}{ c c c c c }
		\toprule
		\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
				\hhline{~----}
			&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
			&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
		\midrule
		1 &1887545 &1887545 &1887545 &2,0734 \\
		2 &2341327 &2341327 &2341327 &2,1451 \\
		3 &3213867 &3213867 &3213867 &\color{red}10,0718 \\
		4 &\color{red}133238294 &\color{red}133238294 &731229 &2,1620 \\
		5 &\color{red}92384021611 &6367983 &\color{red}92384021611 &2,4253 \\
		6 &708335 &708335 &708335 &1,7948 \\
	    \bottomrule
	  \end{tabular}
	\end{table}

@@

	\begin{table}[h!]
	  \centering
	  \tiny
	  \caption{MOS schemes applied to port scan and flood detection}
	  \label{tab:tab4}
	  \begin{tabular}{ c c c c c c c c }
		\toprule
		\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
				\hhline{~------~}
			&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
		\midrule
		Detection of synflood \\(presence of attack) &2 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
		Detection of synflood \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
		\midrule
		Detection of fraggle \\(presence of attack) &1 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
		Detection of fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
		\midrule
		Detection of port scan \\(presence of attack) &1 &1 &\textbf{\color{red}1} &1 &\textbf{\color{red}1} &9 &\textbf{\color{red}1} \\
		Detection of port scan \\(absence of attack) &0 &0 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} \\
		\midrule
		Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} \\
		Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
	    \bottomrule
	  \end{tabular}
	\end{table}

@@

	MOS schemes for probing and flooding attack detection:
	\begin{itemize}
		\item The \textbf{largest eigenvalues visually} indicate the time frame under probing and flooding attack detection;
		\item MOS schemes can estimate it mathematically;
		\begin{itemize}
			\item \textbf{EDC and EFT} estimate \textbf{correctly} the number of attacks $d$;
		\end{itemize}
	\end{itemize}

@@

	\begin{table}[h!]
	  \centering
	  \tiny
	  \caption{Eigen Similarity Analysis for Port Scan Detection}
	  \label{tab:tab5}
	  \begin{tabular}{ c c c c c c }
		\toprule
		\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
				\hhline{~~---~}
				& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
		\midrule
		3 &1 &0.9946 &0.9946 &0.9946 &no \\
		3 &2 &0.9934 &0.9934 &0.9999 &no \\
		3 &3 &0.9912 &0.9912 &0.9999 &no \\
		3 &4 &0.9888 &0.9888 &0.9999 &no \\
		3 &5 &0.9856 &0.9856 &0.9998 &no \\
		3 &6 &0.9840 &0.9840 &0.9999 &no \\
		3 &7 &0.9824 &0.9824 &1.0000 &no \\
		3 &8 &0.9794 &0.9794 &0.9999 &no \\
		3 &9 &0.9673 &0.9673 &0.9926 &no \\
		3 &10 &0.9674 &0.9674 &0.9997 &no \\
		3 &11 &0.9733 &0.9733 &0.9993 &no \\
		3 &12 &0.9702 &0.9702 &0.9993 &no \\
		3 &13 &0.9677 &0.9677 &0.9999 &no \\
		3 &14 &0.9646 &0.9646 &0.9998 &no \\
		3 &15 &\color{red}0.0216 &\color{red}0.0216 &\color{red}0.0276 &\color{red}yes \\
		3 &16 &0.9621 &\color{red}0.0209 &1.0000 &no \\
		3 &17 &0.9611 &\color{red}0.0199 &0.9998 &no \\
		3 &18 &0.9612 &\color{red}0.0191 &0.9999 &no \\
		3 &19 &0.9613 &\color{red}0.0186 &0.9998 &no \\
		3 &20 &0.9638 &\color{red}0.0190 &1.0000 &no \\
	    \bottomrule
	  \end{tabular}
	\end{table}

@@

	\begin{table}[h!]
	  \centering
	  \tiny
	  \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
	  \label{tab:tab8}
	  \begin{tabular}{ c c c c }
		\toprule
		\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
				\hhline{~--~}
				&\textbf{Incremental Individualized} &\textbf{Individual}\\
		\midrule
		80 &0.9999 &0.9999 &no \\
		443 &0.9999 &0.9999 &no \\
		53 &0.9999 &0.9999 &no \\
		21 &0.9999 &0.9997 &\color{red}yes \\
		22 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		23 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		25 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		110 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		143 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		161 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		69 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		123 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		445 &\color{red}0.0298 &0.9997 &\color{red}yes \\
		600 &0.9999 &0.9999 &no \\
		19 &0.9999 &0.9999 &no \\
		67 &0.9999 &0.9999 &no \\
		68 &0.9999 &0.9999 &no \\
	    \bottomrule
	  \end{tabular}
	\end{table}
